---
title: "From Geo2R"
output: html_notebook
---

```{r}
# Version info: R 4.2.2, Biobase 2.58.0, GEOquery 2.66.0, limma 3.54.0
################################################################
#   Differential expression analysis with DESeq2
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
```

```{r}
BiocManager::install("DESeq2")
BiocManager::install("clusterProfiler")
```

```{r}
library(Biobase)
library(GEOquery)
library(limma)
library(DESeq2)
library(clusterProfiler)
```

Because I analyse locally, I need the gset and to set a results directory

```{r}
resfolder <- "GEO2R.GSE218012.results"
if (! file.exists(resfolder) ){
  dir.create(resfolder, showWarnings = FALSE, recursive = FALSE, mode = "0777")
  Sys.chmod(resfolder, mode = "0777", use_umask=TRUE)
}
if (! exists("gset") ){
  gset <- getGEO("GSE218012", destdir = resfolder, GSEMatrix =TRUE)
}
```

Loading Data and loading groups etc.

```{r}
# load counts table from GEO
urld <- "https://www.ncbi.nlm.nih.gov/geo/download/?format=file&type=rnaseq_counts"
path <- paste(urld, "acc=GSE218012", "file=GSE218012_raw_counts_GRCh38.p13_NCBI.tsv.gz", sep="&");
tbl <- as.matrix(data.table::fread(path, header=T, colClasses="integer"), rownames="GeneID")

# load gene annotations 
apath <- paste(urld, "type=rnaseq_counts", "file=Human.GRCh38.p13.annot.tsv.gz", sep="&")
annot <- data.table::fread(apath, header=T, quote="", stringsAsFactors=F, data.table=F)
rownames(annot) <- annot$GeneID
```

```{r}
# sample selection
gsms <- paste0("11111111111111111111111111111111111111111111111111",
        "11111111111111111111111111111111111111111111111111",
        "00000000000000000000000000000000000000000000000000",
        "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")
sml <- strsplit(gsms, split="")[[1]]

# filter out excluded samples (marked as "X")
sel <- which(sml != "X")
sml <- sml[sel]
tbl <- tbl[ ,sel]

# group membership for samples
gs <- factor(sml)
groups <- make.names(c("hEDS","control"))
levels(gs) <- groups
sample_info <- data.frame(Group = gs, row.names = colnames(tbl))

# pre-filter low count genes
# keep genes with at least N counts > 10, where N = size of smallest group
keep <- rowSums( tbl >= 10 ) >= min(table(gs))
tbl <- tbl[keep, ]

ds <- DESeqDataSetFromMatrix(countData=tbl, colData=sample_info, design= ~Group)

ds <- DESeq(ds, test="Wald", sfType="poscount")
```
## Set thresholds for p-value and log2FoldChange for plots
```{r}
log2FCThreshold = 0.5
pValueThreshold = 0.05
```

## Differentially expressed genes

```{r}
# extract results for top genes table
r <- results (ds, contrast=c("Group", groups[1], groups[2]), alpha=0.05, pAdjustMethod ="fdr")

# tT <- r[order(r$padj)[1:250],]
tT <- r[order(r$padj),]
tT <- merge(as.data.frame(tT), annot, by=0, sort=F)

tT <- subset(tT, select=c("EnsemblGeneID","GeneID","padj","pvalue","lfcSE","stat","log2FoldChange","baseMean","Symbol","Description"))
# save all in a file
write.table(tT, file='data/deg.txt', row.names=F, sep=",")
tT
```

## Look only at DEG with p-value <= 0.05 and |log2FoldChange| >= 0.5
```{r}
geneIDs <- subset(tT, select="GeneID", padj<=pValueThreshold& abs(log2FoldChange) >= log2FCThreshold)
print(length(geneIDs$GeneID))
```


```{r}
geneIDs <- subset(tT, select=c("EnsemblGeneID", "GeneID", "Symbol"), padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold)
write.table(geneIDs, file='data/DEG_thresholded.txt', row.names=F, sep=",")
```

## There are some genes without ensemblID, have a closer look there
```{r}
relevantWithoutEnsembl <- subset(tT, select=c("EnsemblGeneID", "GeneID", "Symbol"), padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold & EnsemblGeneID == "")
write.table(relevantWithoutEnsembl, file='data/DEG_thresholded_Ensembl_missing.txt', row.names=F, sep=",")
relevantWithoutEnsembl
```

## Look at amount of up and downregulated genes
```{r}
upregulated = subset(tT, select="Symbol", padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold & log2FoldChange > 0)$Symbol
downregulated = subset(tT, select="Symbol", padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold & log2FoldChange < 0)$Symbol
print(length(upregulated))
print(length(downregulated))
```


## Dispersion Estimates

```{r}
plotDispEsts(ds, main="GSE218012 Dispersion Estimates")
```

## Histogram of p-values

```{r}
# create histogram plot of p-values
hist(r$padj, breaks=seq(0, 1, length = 21), col = "grey", border = "white", 
         xlab = "", ylab = "", main = "GSE218012 Frequencies of padj-values")
```

## Volcano plot

```{r}
# volcano plot
old.pal <- palette(c("#00BFFF", "#FF3030")) # low-hi colors
par(mar=c(4,4,2,1), cex.main=1.5)
plot(r$log2FoldChange, -log10(r$padj), main=paste(groups[1], "vs", groups[2]),
     xlab="log2FC", ylab="-log10(Padj)", pch=20, cex=0.5)
with(subset(r, padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold),
     points(log2FoldChange, -log10(padj), pch=20, col=(sign(log2FoldChange) + 3)/2, cex=1))
legend("bottomleft", title=paste("Padj<", 0.05, sep=""), legend=c("down", "up"), pch=20,col=1:2)
```

## Mean difference (MD) plot

log2 fold change vs. average log2 expression to visualize differentially expressed genes

```{r}
# MD plot
par(mar=c(4,4,2,1), cex.main=1.5)
plot(log10(r$baseMean), r$log2FoldChange, main=paste(groups[1], "vs", groups[2]),
     xlab="log10(mean of normalized counts)", ylab="log2FoldChange", pch=20, cex=0.5)

with(subset(r, padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold),
     points(log10(baseMean), log2FoldChange, pch=20, col=(sign(log2FoldChange) + 3)/2, cex=1))
legend("bottomleft", title=paste("Padj<", 0.05, sep=""), legend=c("down", "up"), pch=20,col=1:2)
abline(h=0)
palette(old.pal) # restore palette
```

## General expression data visualization

-   distribution of the values of the selected samples

```{r}
################################################################
#   General expression data visualization
dat <- log10(counts(ds, normalized = T) + 1) # extract normalized counts

# box-and-whisker plot
lbl <- "log10(raw counts + 1)"
dev.new(width=3+ncol(gset)/6, height=5)
ord <- order(gs)  # order samples by group
palette(c("#1B9E77", "#7570B3", "#E7298A", "#E6AB02", "#D95F02",
          "#66A61E", "#A6761D", "#B32424", "#B324B3", "#666666"))
par(mar=c(7,4,2,1))
boxplot(dat[,ord], boxwex=0.6, notch=T, main="GSE218012", ylab="lg(norm.counts)", outline=F, las=2, col=gs[ord])
legend("topleft", groups, fill=palette(), bty="n")
```

## Uniform Manifold Approximation and Projection (UMAP) plot

Dimensionality reduction to visualize how samples are related to each other

```{r}
# UMAP plot (multi-dimensional scaling)
library(umap)
dat <- dat[!duplicated(dat), ] # first remove duplicates
par(mar=c(3,3,2,6), xpd=TRUE, cex.main=1.5)
ump <- umap(t(dat), n_neighbors = 15, random_state = 123)
plot(ump$layout, main="UMAP plot, nbrs=15", xlab="", ylab="", col=gs, pch=20, cex=1.5)
legend("topright", inset=c(-0.15,0), legend=groups, pch=20,
       col=1:length(groups), title="Group", pt.cex=1.5)
```

For the final network, where we query the known EDS genes of other types, it is interesting to retrieve the pValues and log2FoldChange for those as well, even if there are not differentially expressed under the chosen conditions.

```{r}
# extract all results
r_eds <- results (ds, contrast=c("Group", groups[1], groups[2]), alpha=0.99999, lfcThreshold=0, altHypothesis ="greaterAbs", pAdjustMethod ="fdr")

tT_eds <- r_eds[order(r_eds$padj),]
tT_eds <- merge(as.data.frame(tT_eds), annot, by=0, sort=F)

eds_symbols = c("ADAMTS2",
"AEBP1",
"B3GALT6",
"B4GALT7",
"C1R",
"C1S",
"CCBE1",
"CHST14",
"COL12A1",
"COL1A1",
"COL1A2",
"COL3A1",
"COL5A1",
"COL5A2",
"DSE",
"FKBP14",
"PLOD1",
"PRDM5",
"SLC39A13",
"SLC6A5",
"TNXB",
"ZNF469"
)

tT_eds <- subset(tT_eds, Symbol %in% eds_symbols)

tT_eds <- subset(tT_eds, select=c("GeneID","padj","pvalue","lfcSE","stat","log2FoldChange","baseMean","Symbol","Description"))
#write.table(tT, file=stdout(), row.names=F, sep="\t")
# tT <- subset(tT, select=c("GeneID","padj","pvalue","lfcSE","stat","log2FoldChange","baseMean","Symbol","Description"), padj<pValueThreshold & abs(log2FCThreshold) >= log2FCThreshold)
write.table(tT_eds, file='data/eds_expression.txt', row.names=F, sep=",")
tT_eds
```