---
title: "Community Clustering"
output: html_notebook
---

```{r}
BiocManager::install(c("igraph", "RCy3", "clusterProfiler", "org.Hs.eg.db"))
```


```{r}
library(RCy3)
library(igraph)
library(clusterProfiler)
library(org.Hs.eg.db)

```

```{r}

cytoscapePing ()
```

```{r}
if(FALSE) {
networkSuid = getNetworkSuid()
print(networkSuid)
}
```

```{r}
commandsHelp("help cluster")
```

```{r}
# minimum number of nodes to create a new network out of the cluster
nodeThreshold <- 15
```

# Clustering in cytoscape
```{r}
if(FALSE) {
  clustermaker = paste("cluster glay network=SUID:", networkSuid, sep="")
  res <- commandsGET(clustermaker)
  num <- as.numeric(gsub("Clusters: ", "", res[1]))
  print(paste("Found", num, "clusters"))
  
  clusterNumber <- 0
  # for each cluster
  for(i in 1:num) {
    setCurrentNetwork(network=networkSuid)
    clearSelection(network=networkSuid)
    currentCluster <- i
    selectedNodes <- selectNodes(network=networkSuid, currentCluster, by.col="__glayCluster") 
    if(length(selectedNodes$nodes) > nodeThreshold) {
      clusterNumber <- clusterNumber + 1
      # sub network creation
      subnetwork_suid <- createSubnetwork(nodes="selected", network=networkSuid)
      table <- getTableColumns(table="node",columns = c("display name", "Ensembl", "Entrez Gene"), network = as.numeric(subnetwork_suid))
      # save network
      exportNetwork(paste("data/networks/glay/glay_cluster_", clusterNumber, ".sif", sep=""), network=subnetwork_suid)
      res <- clusterProfiler::enrichGO(
        gene = table$`display name`,
        OrgDb = org.Hs.eg.db,
        keyType = "SYMBOL",
        ont = "BP",
        pAdjustMethod ="BH",
        pvalueCutoff = 0.05,
        qvalueCutoff = 0.05,
        readable = TRUE
        )
      res.df <- as.data.frame(res)
      write.table(res.df, file=paste('data/enrichment/biological-process/glay_cluster', clusterNumber, '_bp.txt', sep=""), row.names=F, sep=",")
      write.table(table, file=paste("data/networks/glay/glay_cluster", clusterNumber,"_nodes.txt", sep=""), row.names=F, sep=",")
      assign(paste("glay_Cluster_", clusterNumber, sep=""), res.df)
      barplot(res, showCategory = 20, title = paste("Biological process glay cluster", clusterNumber))
    }
  }
}
```
## Check whether the MCODE cluster 1 ECM genes are in the cluster

```{r}
ecmGenes <- c("COL10A1", "COL15A1", "PCOLCE", "COL5A3", "COL18A1", "COL21A1", "COL27A1")
tableCluster1 <- read.csv(paste("data/networks/glay/glay_cluster", 1,"_nodes.txt", sep=""), header=TRUE)
for(gene in ecmGenes) {
  cat(paste("ECM gene", gene ,"in glay cluster 1?", gene %in% tableCluster1$`display.name`, "\n"))
}
```
## Check whether the chromatin genes are in a cluster (I saw some in cluster 5)
There are none in community cluster 1
```{r}
chromatinGenes <- c(
  "H1-3", "H1-4", "H1-5", "H2AC8", "H2AC7", "H2AC13", "H2AC16", "H2AC4", 
  "H2AC17", "H2BC8", "H2BC15", "H2BC14", "H2BC7", "H2BC6", "H2BC9", "H2BC10", 
  "H2BC4", "H2BC21", "H3C1", "H3C4", "H3C3", "H3C6", "H3C11", "H3C8", "H3C12", 
  "H3C2", "H4C1", "H4C6", "H4C12", "H4C11", "H4C3", "H4C8", "H4C13", "H3C7", 
  "H2AC11", "H2AC21")
tableCluster5 <- read.csv(paste("data/networks/glay/glay_cluster", 5,"_nodes.txt", sep=""), header=TRUE)
for(gene in chromatinGenes) {
  cat(paste("chromatin gene", gene ,"in glay cluster 5?", gene %in% tableCluster5$`display.name`, "\n"))
}
```



## Enrichment for Biological Process
```{r}
enrichmentBiologicalProcess <- function(i) {
  tableCluster <- read.csv(paste("data/networks/glay/glay_cluster", i,"_nodes.txt", sep=""), header=TRUE)
  res <- clusterProfiler::enrichGO(
      gene = tableCluster$`display.name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "BP",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  res.df <- as.data.frame(res)
  cat(capture.output(head(res), sep="\n"))
  return(barplot(res, showCategory = 8, title = paste("Biological Process community cluster", i), font.size = 11))
}
```

```{r}
enrichmentBiologicalProcess(1, fig.dpi=300, fig.height=4, fig.width=5)
```
```{r}
enrichmentBiologicalProcess(2)
```

```{r}
enrichmentBiologicalProcess(3)
```

```{r, fig.dpi=300, fig.height=4, fig.width=6}
enrichmentBiologicalProcess(4)
```

```{r, fig.dpi=300, fig.height=4, fig.width=6}
enrichmentBiologicalProcess(5)
```
# Only 15 genes so probably not helpful (Why 15, it should be 18???)
```{r}
enrichmentBiologicalProcess(6)
```

## Analyse for cellular component as well
```{r}
enrichmentCellularComponent <- function(i) {
  tableCluster <- read.csv(paste("data/networks/glay/glay_cluster", i,"_nodes.txt", sep=""), header=TRUE)
  res <- clusterProfiler::enrichGO(
      gene = tableCluster$`display.name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "CC",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  res.df <- as.data.frame(res)
  cat(capture.output(head(res), sep="\n"))
  barplot(res, showCategory = 10, title = paste("Cellular Component glay cluster", i), font.size = 11)
}
```

```{r}
enrichmentCellularComponent(1)
```
```{r}
enrichmentCellularComponent(2)
```

```{r}
enrichmentCellularComponent(3)
```

```{r}
enrichmentCellularComponent(4)
```

```{r}
enrichmentCellularComponent(5)
```

```{r}
enrichmentCellularComponent(6)
```
# Do molecular function for the 3 small networks (2, 3 and 6)

```{r}
enrichmentMolecularFunction <- function(i) {
  tableCluster <- read.csv(paste("data/networks/glay/glay_cluster", i,"_nodes.txt", sep=""), header=TRUE)
  res <- clusterProfiler::enrichGO(
      gene = tableCluster$`display.name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  res.df <- as.data.frame(res)
  cat(capture.output(head(res), sep="\n"))
  barplot(res, showCategory = 10, title = paste("Molecular Function glay cluster", i), font.size = 11)
}
```

```{r}
enrichmentMolecularFunction(2)
```

```{r}
enrichmentMolecularFunction(3)
```

```{r}
enrichmentMolecularFunction(6)
```
# Pathway analysis
Get kegg code for homo sapiens
```{r}
human <- search_kegg_organism('Homo sapiens', by='scientific_name')
human <- human$kegg_code[1]
```

```{r}
enrichKEGGForCluster <- function(i) {
  tableCluster <- read.csv(paste("data/networks/glay/glay_cluster", i,"_nodes.txt", sep=""), header=TRUE)
  kk <- enrichKEGG(
    gene= tableCluster$`Entrez.Gene`,
    organism= human,
  )
head(kk, 15)
}
```

```{r}
enrichKEGGForCluster(1)
```

```{r}
enrichKEGGForCluster(4)
```

```{r}
enrichKEGGForCluster(5)
```

# WikiPathways
```{r}
get_wp_data <- function(organism, output = "data.frame") {
    read.gmt.wp('wikipathways-20231010-gmt-Homo_sapiens.gmt.txt', output = output)
}

prepare_WP_data <- function(organism) {
    wp2gene <- get_wp_data(organism)
    ##TERM2GENE
    wpid2gene <- wp2gene %>% dplyr::select(.data$wpid, .data$gene) 
    ##TERM2NAME
    wpid2name <- wp2gene %>% dplyr::select(.data$wpid, .data$name) 
    list(WPID2GENE = wpid2gene,
         WPID2NAME = wpid2name)
}

# enrichWP(geneIDs, organism="Homo sapiens")
wiki_pathway <- read.gmt.wp('wikipathways-20231010-gmt-Homo_sapiens.gmt.txt')

enrichWP <- function(gene, organism, ...) {
    wpdata <- prepare_WP_data(organism)
    res <- enricher(gene,
                    TERM2GENE = wpdata$WPID2GENE,
                    TERM2NAME = wpdata$WPID2NAME,
                    ...)
    if (is.null(res)) return(res)

    res@ontology <- "WikiPathways"
    res@organism <- organism
    res@keytype <-  "ENTREZID"

    return(res)
}
```

```{r}
enrichWPForCluster <- function(i) {
  tableCluster <- read.csv(paste("data/networks/glay/glay_cluster", i,"_nodes.txt", sep=""), header=TRUE)
  wp <- enrichWP(
    gene= tableCluster$`Entrez.Gene`,
    organism= human,
  )
  head(wp, 15)
}
```


```{r}
enrichWPForCluster(1)
```

```{r}
enrichWPForCluster(4)
```

```{r}
enrichWPForCluster(5)
```

