a
---
title: "Enrichment"
output: html_notebook
---

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(BiocManager)
BiocManager::install("clusterProfiler")
BiocManager::install("org.Hs.eg.db")
BiocManager::install("pathview")
BiocManager::install("ReactomePA")
BiocManager::install("KEGGgraph")
```

```{r}
library("clusterProfiler")
library("pathview")
library("ReactomePA")
```

```{r}
df = read.csv("data/deg.txt", header=TRUE)
df
```
```{r}
pValueThreshold <- 0.05
log2FCThreshold <- 0.5
geneList <- subset(df, select=c("GeneID", "log2FoldChange"), padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold)
foldChange <- geneList$log2FoldChange
geneListOrdered <- setNames(foldChange, as.character(geneList$GeneID))
geneListOrdered <- sort(geneListOrdered, decreasing=T)
```
## Gene Ontology
### Molecular function
#### Go classification
```{r}
geneIDs <- as.character(subset(df, select="GeneID", padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold)$GeneID)
ggo_mf <- groupGO(
  gene=geneIDs,
  "org.Hs.eg.db",
  ont="MF",
  level=3,
  readable=TRUE
)
ggo_mf <- ggo_mf[order(ggo_mf$Count, decreasing=T),]
head(ggo_mf, 25)
```
#### Overrepresentation
```{r}
ego_mf <- enrichGO(gene=geneIDs, "org.Hs.eg.db", ont = "MF", readable = T, pvalueCutoff = 0.05)
head(ego_mf,20)
```
```{r}
barplot(ego_mf, showCategory = 20, title = "Molecular Function")
```

```{r}
goplot(ego_mf)
```
#### GSEO
```{r}
enrichmentGo_mf <- gseGO(geneListOrdered, "org.Hs.eg.db", ont="MF")
head(enrichmentGo_mf, 20)
```
### Cellular component
#### Go classification
```{r}
ggo_cc <- groupGO(
  gene=geneIDs,
  "org.Hs.eg.db",
  ont="CC",
  level=3,
  readable=TRUE
)
ggo_cc <- ggo_cc[order(ggo_cc$Count, decreasing=T),]
head(ggo_cc, 25)
```
#### Overrepresentation
```{r}
ego_cc <- enrichGO(gene=geneIDs, "org.Hs.eg.db", ont = "CC", readable = T, pvalueCutoff = 0.05)
head(ego_cc,20)
```
GO-Term without description: OBSOLETE. A protein complex that plays a role in the process of DNA packaging.

###### Look which genes are in collagen-containing extracellular matrix
```{r}
geneSymbols_ego_cc <- bitr(ego_cc@geneSets$`GO:0062023`, fromType="ENTREZID", toType="SYMBOL", OrgDb="org.Hs.eg.db")
cat(paste(geneSymbols_ego_cc$SYMBOL, collapse=", "))
```

```{r}
barplot(ego_cc, showCategory = 15, title = "Cellular Component", font.size = 8)
```
```{r}
dotplot(ego_cc, showCategory = 15, title = "Differentially expressed genes - Cellular Component", font.size = 8)
```

```{r}
goplot(ego_cc)
```

#### Go enrichment
```{r}
enrichmentGo_cc <- gseGO(geneListOrdered, "org.Hs.eg.db", ont="CC")
head(enrichmentGo_cc, 20)
```
### Biological process

#### Go classification
```{r}
ggo_bp <- groupGO(
  gene=geneIDs,
  "org.Hs.eg.db",
  ont="BP",
  level=3,
  readable=TRUE
)
ggo_bp <- ggo_bp[order(ggo_bp$Count, decreasing=T),]
head(ggo_bp, 25)
```
#### Overrepresentation
```{r}
ego_bp <- enrichGO(gene=geneIDs, "org.Hs.eg.db", ont = "BP", readable = T, pvalueCutoff = 0.05)
head(ego_bp,20)
```
```{r}
barplot(ego_bp, showCategory = 10, title = "Biological Process", font.size = 9)
```

```{r}
dotplot(ego_bp, showCategory = 15, title = "Biological Process", font.size = 8)
```

```{r}
goplot(ego_bp)
```

#### Go enrichment
```{r}
enrichmentGo_bp <- gseGO(geneListOrdered, "org.Hs.eg.db", ont="BP")
head(enrichmentGo_bp, 20)
```
## KEGG

Get kegg code for homo sapiens
```{r}
human <- search_kegg_organism('Homo sapiens', by='scientific_name')
human <- human$kegg_code[1]
```

### KEGG pathway over-representation analysis
```{r}
kk <- enrichKEGG(
  gene= geneIDs,
  organism= human
)
head(kk, 15)
```
### KEGG pathway gene set enrichment analysis
```{r}
kkGSE <- gseKEGG(
  gene=geneListOrdered,
  organism=human
)
head(kkGSE, 15)
```
### KEGG module over-representation analysis

```{r}
mkk <- enrichMKEGG(gene = geneIDs,
                   organism = 'hsa')
print(length(mkk$ID))
head(mkk)
```
### KEGG module gene set enrichment analysis
```{r}
gseMKK <- gseMKEGG(gene = geneListOrdered,
                   organism = 'hsa')
head(gseMKK) 
```
```{r}

# hsa04110 <- pathview(gene.data=geneListOrdered, pathway.id="hsa04110", species=human) # cell cycle
```
## WikiPathways

```{r}
get_wp_data <- function(organism, output = "data.frame") {
    read.gmt.wp('wikipathways-20231010-gmt-Homo_sapiens.gmt.txt', output = output)
}

prepare_WP_data <- function(organism) {
    wp2gene <- get_wp_data(organism)
    ##TERM2GENE
    wpid2gene <- wp2gene %>% dplyr::select(.data$wpid, .data$gene) 
    ##TERM2NAME
    wpid2name <- wp2gene %>% dplyr::select(.data$wpid, .data$name) 
    list(WPID2GENE = wpid2gene,
         WPID2NAME = wpid2name)
}

# enrichWP(geneIDs, organism="Homo sapiens")
wiki_pathway <- read.gmt.wp('wikipathways-20231010-gmt-Homo_sapiens.gmt.txt')

enrichWP <- function(gene, organism, ...) {
    wpdata <- prepare_WP_data(organism)
    res <- enricher(gene,
                    TERM2GENE = wpdata$WPID2GENE,
                    TERM2NAME = wpdata$WPID2NAME,
                    ...)
    if (is.null(res)) return(res)

    res@ontology <- "WikiPathways"
    res@organism <- organism
    res@keytype <-  "ENTREZID"

    return(res)
}

wp_enrichment <- enrichWP(geneIDs, "Homo sapiens", pvalueCutoff = 0.05)
head(wp_enrichment)
```
### WikiPathways gene set enrichment analysis
```{r}
gseWP <- function(geneList, organism, ...) {
    wpdata <- prepare_WP_data(organism)
    res <- GSEA(geneList,
                TERM2GENE = wpdata$WPID2GENE,
                TERM2NAME = wpdata$WPID2NAME,
                ...)

    if (is.null(res)) return(res)

    res@setType <- "WikiPathways"
    res@organism <- organism
    res@keytype <-  "ENTREZID"

    return(res)
}
gse_wikipathways <- gseWP(geneListOrdered, organism = "Homo sapiens")
head(gse_wikipathways, 15)
```
## Reactome
### Overrepresentation
```{r}
x <- enrichPathway(gene=geneIDs, pvalueCutoff = 0.05, readable=TRUE)
head(x)
```

### Gene set enrichment
```{r}
y <- gsePathway(geneListOrdered)
print(length(y$ID))
head(y,20)
```

