---
title: "Heat Diffusion"
output: html_notebook
---

```{r}
df = read.csv("deg.txt", header=TRUE)
```

```{r}
BiocManager::install(c("RCy3"))
```

```{r}
library(RCy3)
```

```{r}
clearSelection()
copyVisualStyle("default", "diffusion")
setVisualStyle("diffusion")
deleteStyleMapping("diffusion","NODE_FILL_COLOR")
```

# Select network

```{r}
networkSuid <- getNetworkSuid()
print(networkSuid)
```

# select start nodes

```{r}
edsGeneSymbols = c("ADAMTS2", "AEBP1", "B3GALT6", "B4GALT7", "C1R", "C1S", "CCBE1", "CHST14", "COL12A1", "COL1A1", "COL1A2", "COL3A1","COL5A1", "COL5A2", "DSE", "FKBP14", "PLOD1", "PRDM5", "SLC39A13", "SLC6A5", "TNXB", "ZNF469")
```

```{r}
knownEDS.table <- getTableColumns('node','knownEDS', network=networkSuid)
edsNodes <- row.names(knownEDS.table)[knownEDS.table$knownEDS == TRUE]
cat(paste(edsNodes, collapse=", "))
cat(paste("\n", length(edsNodes), "nodes related to other EDS types"))
selectNodes(edsNodes)
setNodeBorderColorBypass(edsNodes,"#FF55AA")
setNodeBorderWidthBypass(edsNodes,5)
```

# Diffusion

```{r}
diffusionBasic()
clearSelection()
setNodeColorMapping("diffusion_output_heat", c(0,0.5), c("#FFFFFF", "#006699"), style="diffusion")
```

# select nodes with highest ranks and look at diffuasion heat before to get a feeling for the amount of hot nodes

```{r}
networkSuidDiffusion <- getNetworkSuid()
cat(networkSuidDiffusion)
nodesAfterDiffusion <- getTableColumns('node', c('knownEDS', "display name", "Entrez Gene", "diffusion_output_heat", "diffusion_input", "diffusion_output_rank"), network=networkSuidDiffusion)
head(nodesAfterDiffusion[order(nodesAfterDiffusion$diffusion_output_rank),])
```

```{r}
hist(nodesAfterDiffusion$diffusion_output_heat, breaks = "sturges")
```

```{r}
cat(length(subset(nodesAfterDiffusion, diffusion_output_heat > 0.3)$diffusion_output_rank))
cat("\n")
cat(length(subset(nodesAfterDiffusion, diffusion_output_heat > 0.2)$diffusion_output_rank))
cat("\n")
cat(length(subset(nodesAfterDiffusion, diffusion_output_heat > 0.1)$diffusion_output_rank))
```

```{r}
subset_without_eds <- subset(nodesAfterDiffusion, knownEDS == FALSE & diffusion_output_heat > 0.1)
hotGenes <- subset_without_eds[order(subset_without_eds$diffusion_output_rank),]
cat(paste("Amount of hot genes", length(hotGenes$diffusion_output_rank)))
head(hotGenes, 15)
```

# Enrichment on hot genes (compare to mcode cluster 3 as well)

```{r}
plotMFEnrichment <- function(genes) {
  res <- clusterProfiler::enrichGO(
      gene = genes,
      OrgDb = org.Hs.eg.db,
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  resr.df <- as.data.frame(res)
  cat(capture.output(head(res), sep="\n"))
  barplot(res, showCategory = 8, title = paste("Molecular Function", sep=""), font.size=10)
  #return(resCluster)
}
```

```{r}
plotMFEnrichment(hotGenes$`Entrez Gene`)
```

```{r}
hotGenes$`Entrez Gene`
```

# Check for overlap with mcode cluster 3 genes

```{r}
mcode_cluster_3_genes = read.csv(paste("data/networks/MCODE/mcode_cluster_", 3,"_nodes.txt", sep=""), header=TRUE)$`display.name`
cat(paste(mcode_cluster_3_genes, collapse=", "))
cat("\n")
cat(paste(hotGenes$`display name`, collapse=", "))
cat("\n")
cat("intersect\n")
intersect(mcode_cluster_3_genes, hotGenes$`display name`)
```

No intersect besides EDS genes but still similar molecular function. Are the MCODE cluster 3 genes even in community cluster 1?#

# the heat decays pretty quick for such a large graph, so repeat with higher spread

```{r}
networkSuid <- getNetworkSuid()
print(networkSuid)
```

```{r}
clearSelection()
copyVisualStyle("default", "diffusion")
setVisualStyle("diffusion")
deleteStyleMapping("diffusion","NODE_FILL_COLOR")
```

## select start nodes

```{r}
edsGeneSymbols = c("ADAMTS2", "AEBP1", "B3GALT6", "B4GALT7", "C1R", "C1S", "CCBE1", "CHST14", "COL12A1", "COL1A1", "COL1A2", "COL3A1","COL5A1", "COL5A2", "DSE", "FKBP14", "PLOD1", "PRDM5", "SLC39A13", "SLC6A5", "TNXB", "ZNF469")

knownEDS.table <- getTableColumns('node','knownEDS', network=networkSuid)
edsNodes <- row.names(knownEDS.table)[knownEDS.table$knownEDS == TRUE]
cat(paste(edsNodes, collapse=", "))
cat(paste("\n", length(edsNodes), "nodes related to other EDS types"))
selectNodes(edsNodes)
setNodeBorderColorBypass(edsNodes,"#FF55AA")
setNodeBorderWidthBypass(edsNodes,5)
```

## Diffusion

```{r}
diffusionAdvanced(time=0.3)
clearSelection()
setNodeColorMapping("diffusion_output_heat", c(0,0.5), c("#FFFFFF", "#006699"), style="diffusion")
```

```{r}
networkSuidDiffusion03 <- getNetworkSuid()
cat(networkSuidDiffusion03)
nodesAfterDiffusion03 <- getTableColumns('node', c('knownEDS', "display name", "Entrez Gene", "diffusion_output_heat", "diffusion_input", "diffusion_output_rank", "knownEDS"), network=networkSuidDiffusion03)
head(nodesAfterDiffusion03[order(nodesAfterDiffusion03$diffusion_output_rank),])
```

```{r}
hist(nodesAfterDiffusion03$diffusion_output_heat, breaks = "sturges")
```

```{r}
cat(length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.3)$diffusion_output_rank))
cat("\n")
cat(length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.2)$diffusion_output_rank))
cat("\n")
cat(length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.1)$diffusion_output_rank))
cat("\n")
cat(length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.05)$diffusion_output_rank))
```

```{r}
subset_without_eds03 <- subset(nodesAfterDiffusion03, knownEDS == FALSE & diffusion_output_heat > 0.1)
hotGenes03 <- subset_without_eds03[order(subset_without_eds03$diffusion_output_rank),]
cat(paste("Amount of hot genes", length(hotGenes03$diffusion_output_rank)))
head(hotGenes03, 15)
```

```{r}
plotMFEnrichment(hotGenes03$`Entrez Gene`)
```
# Check for overlap with mcode cluster 3 genes

```{r}
mcode_cluster_3_genes = read.csv(paste("data/networks/MCODE/mcode_cluster_", 3,"_nodes.txt", sep=""), header=TRUE)$`display.name`
cat(paste(mcode_cluster_3_genes, collapse=", "))
cat("\n")
cat(paste(hotGenes03$`display name`, collapse=", "))
cat("\n")
cat("intersect\n")
intersectingGenes <-intersect(mcode_cluster_3_genes, hotGenes03$`display name`)
cat(paste(intersectingGenes, collapse=", "))
cat("\n setdiff \n")
nonIntersecting <- c()
for(gene in hotGenes03$`display name`) {
  if (!(gene %in% edsGeneSymbols) & !(gene %in% intersectingGenes)) {
    nonIntersecting <- c(nonIntersecting, gene)
  }
}
cat("\n")
cat(paste(nonIntersecting, collapse=", "))
```
## Which are the hottest genes that are no EDS genes
```{r}
mcode_cluster_3_genes = read.csv(paste("data/networks/MCODE/mcode_cluster_", 3,"_nodes.txt", sep=""), header=TRUE)$`display.name`
cat(paste("Genes with heat > 0.3:", length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.3 & knownEDS == FALSE)$diffusion_output_rank)))
cat("\n")
cat(paste(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.3 & knownEDS == FALSE)$`display name`, collapse=", "))
cat("\n ---- \n")
cat(paste("Genes with heat > 0.2:", length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.2 & knownEDS == FALSE)$diffusion_output_rank)))
cat("\n")
cat(paste(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.2 & knownEDS == FALSE)$`display name`, collapse=", "))
cat("\n ---- \n")
cat(paste("Genes with heat > 0.1:", length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.1 & knownEDS == FALSE)$diffusion_output_rank)))
cat("\n")
nonEDSHot03 <- subset(nodesAfterDiffusion03, diffusion_output_heat > 0.1 & knownEDS == FALSE)
cat(paste(nonEDSHot03$`display name`, collapse=", "))
cat("\nGenes without the ones in EDS MCODE cluster\n")
nonEDSHot03_notInMCODE <- c()
for(gene in nonEDSHot03$`display name`) {
  if (!gene %in% mcode_cluster_3_genes) {
    nonEDSHot03_notInMCODE <- c(nonEDSHot03_notInMCODE, gene)
  }
}
cat(paste(nonEDSHot03_notInMCODE, collapse=", "))
cat(paste(" (in total", length(nonEDSHot03_notInMCODE), ")"))
cat("\n ---- \n")
cat(paste("Genes with heat > 0.05:", length(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.05 & knownEDS == FALSE)$diffusion_output_rank)))
cat("\n")
cat(paste(subset(nodesAfterDiffusion03, diffusion_output_heat > 0.05 & knownEDS == FALSE)$`display name`, collapse=", "))
cat("\n ---- \n")
```
# Look at hot gene go terms
```{r}

resHotNotEDSMF <- clusterProfiler::groupGO(
      gene = nonEDSHot03_notInMCODE,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      level=3,
      readable = TRUE
      )
resHotNotEDSMF <- resHotNotEDSMF[order(resHotNotEDSMF$Count, decreasing=T),]
head(resHotNotEDSMF, 20)
```

```{r}
resHotNotEDSCC <- clusterProfiler::groupGO(
      gene = nonEDSHot03_notInMCODE,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "CC",
      level=3,
      readable = TRUE
      )
resHotNotEDSCC <- resHotNotEDSCC[order(resHotNotEDSCC$Count, decreasing=T),]
head(resHotNotEDSCC, 20)
```

```{r}
resHotNotEDSBP <- clusterProfiler::groupGO(
      gene = nonEDSHot03_notInMCODE,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "BP",
      level=3,
      readable = TRUE
      )
resHotNotEDSBP <- resHotNotEDSBP[order(resHotNotEDSBP$Count, decreasing=T),]
head(resHotNotEDSBP, 20)
```
```{r}
resAll <- clusterProfiler::enrichGO(
      gene = nonEDSHot03_notInMCODE,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "all",
      readable = TRUE
      )
resAll.df <- as.data.frame(resAll)
head(resAll)
```
