---
title: "MCODE Clustering"
output: html_notebook
fig_width: 4 
fig_height: 3
---

```{r}
BiocManager::install(c("igraph", "RCy3", "clusterProfiler", "org.Hs.eg.db"))
```
# Get libraries
```{r}
library(RCy3)
library(igraph)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ggplot2)
```

```{r}
cytoscapePing ()
```

```{r}
networkSuid = getNetworkSuid()
print(networkSuid)
```

```{r}
commandsHelp("help cluster")
```
```{r}
edsGeneSymbols = c("ADAMTS2", "AEBP1", "B3GALT6", "B4GALT7", "C1R", "C1S", "CCBE1", "CHST14", "COL12A1", "COL1A1", "COL1A2", "COL3A1","COL5A1", "COL5A2", "DSE", "FKBP14", "PLOD1", "PRDM5", "SLC39A13", "SLC6A5", "TNXB", "ZNF469")
```


```{r}
# minimum number of nodes to create a new network out of the cluster
nodeThreshold <- 15
```
# Perform clustering and save clusters
```{r}
clustermaker = paste("cluster mcode network=SUID:", networkSuid, sep="")
res <- commandsGET(clustermaker)
num <- as.numeric(gsub("Clusters: ", "", res[1]))
print(paste("Found", num, "clusters"))

clusterNumber = 0
# for each cluster
for(i in 1:num) {
  setCurrentNetwork(network=networkSuid)
  clearSelection(network=networkSuid)
  currentCluster <- i
  selectedNodes <- selectNodes(network=networkSuid, currentCluster, by.col="__mcodeCluster") 
  print(paste("Cluster", i, "has", length(selectedNodes$nodes), "nodes"))
  if(length(selectedNodes$nodes) > nodeThreshold) {
    # sub network creation
    # cluster numbers of clusters with #nodes > nodeThreshold
    clusterNumber = clusterNumber + 1
    
    # export subnetwork
    subnetwork_suid <- createSubnetwork(nodes="selected", network=networkSuid)
    exportNetwork(paste("data/networks/MCODE/mcode_cluster_", clusterNumber, ".sif", sep=""), network=subnetwork_suid)
    
    # export nodes of subnetwork
    table <- getTableColumns(table="node",columns = c("display name", "Ensembl", "Entrez Gene"), network = as.numeric(subnetwork_suid))
    write.table(table, file=paste("data/networks/MCODE/mcode_cluster_", clusterNumber,"_nodes.txt", sep=""), row.names=F, sep=",")
    
    res <- clusterProfiler::enrichGO(
      gene = table$`display name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )

    res.df <- as.data.frame(res)
    
    # save enrichment results
    write.table(res.df, file=paste('data/enrichment/molecular-function/mcode_cluster', clusterNumber, '_go.txt', sep=""), row.names=F, sep=",")
    
    assign(paste("MCODE_Cluster_", clusterNumber, sep=""), res.df)
    # plot enrichment results
    barplot(res, showCategory = 10, title = paste("Molecular Function MCODE cluster", clusterNumber))
  }
}
```

```{r}
plotClusterEnrichment <- function(i, dot=FALSE, returnResults=FALSE) {
  tableCluster <- read.csv(paste("data/networks/MCODE/mcode_cluster_", i,"_nodes.txt", sep=""), header=TRUE)
  resCluster <- clusterProfiler::enrichGO(
      gene = tableCluster$`display.name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  resCluster.df <- as.data.frame(resCluster)
  cat(capture.output(head(resCluster), sep="\n"))
  if(returnResults){
    return(resCluster)
  } else {
    if(dot) {
    return(dotplot(resCluster, showCategory = 8, title = paste("MCODE cluster ",i, sep=""), font.size=11))
  } else {
    return(barplot(resCluster, showCategory = 8, title = paste("MCODE cluster ",i, sep=""), font.size=11))
  }
  }
}
```


```{r}
plotClusterEnrichment(1)
```
```{r, fig.dpi=300, fig.height=4, fig.width=5}
plotClusterEnrichment(1, TRUE, FALSE)
```


```{r}
plotClusterEnrichment(2)
```
```{r, fig.dpi=300, fig.height=4, fig.width=5}
plotClusterEnrichment(2, TRUE, FALSE)
```

```{r}
plotClusterEnrichment(3)
```

```{r}
plotClusterEnrichment(3, TRUE)
```
Look if only the known EDS genes are in the extracellular matrix molecular function or whether the other genes are there as well (or rather) which ones

##### First for GO:0030020 - extracellular matrix structural constituent conferring tensile strength
```{r}
resCluster3 <- plotClusterEnrichment(3, returnResults=TRUE)
extracellular_tensile_strength_genes <- resCluster3@geneSets$`GO:0030020`
extracellular_tensile_strength_nonEDS <- c()
extracellular_tensile_strength_EDS <- c()
for(gene in extracellular_tensile_strength_genes) {
  if((!(gene %in% edsGeneSymbols)) & (gene %in% resCluster3@gene)) {
    extracellular_tensile_strength_nonEDS <- c(extracellular_tensile_strength_nonEDS, gene)
  }
  else {
    if(gene %in% edsGeneSymbols & (gene %in% resCluster3@gene)) {
      extracellular_tensile_strength_EDS <- c(extracellular_tensile_strength_EDS, gene)
    }
  }
}
edsGenesString <- paste(extracellular_tensile_strength_EDS, collapse = ", ")
nonEDSGenesString <- paste(extracellular_tensile_strength_nonEDS, collapse=", ")
cat(paste("EDS genes in extracellular matrix structural constituent conferring tensile strength", edsGenesString))
cat("\n")
cat(paste("In total", length(extracellular_tensile_strength_EDS), "EDS genes of overall", length(extracellular_tensile_strength_EDS)+length(extracellular_tensile_strength_nonEDS)))
cat("\n")
cat(paste("Non-EDS genes: ", nonEDSGenesString))
```
##### Then for GO:0005201 - extracellular matrix structural constituent
```{r}
extracellular_structure_genes <- resCluster3@geneSets$`GO:0005201`
extracellular_structure_nonEDS <- c()
extracellular_structure_EDS <- c()
for(gene in extracellular_structure_genes) {
  if((!(gene %in% edsGeneSymbols)) & (gene %in% resCluster3@gene)) {
    extracellular_structure_nonEDS <- c(extracellular_structure_nonEDS, gene)
  }
  else {
    if(gene %in% edsGeneSymbols & (gene %in% resCluster3@gene)) {
      extracellular_structure_EDS <- c(extracellular_structure_EDS, gene)
    }
  }
}
edsGenesString <- paste(extracellular_structure_EDS, collapse = ", ")
nonEDSGenesString <- paste(extracellular_structure_nonEDS, collapse=", ")
cat(paste("EDS genes in extracellular matrix structural constituent", edsGenesString))
cat("\n")
cat(paste("In total", length(extracellular_structure_EDS), "EDS genes of overall", length(extracellular_structure_EDS)+length(extracellular_structure_nonEDS)))
cat("\n")
cat(paste("Non-EDS genes: ", nonEDSGenesString))
```
# look at chromatin enrichment
```{r}
resCluster2 <- plotClusterEnrichment(i=2, returnResults = TRUE)
chromatin_genes <- c()
for(gene in resCluster2@geneSets$`GO:0030527`) {
  if(gene %in% resCluster2@gene) {
    chromatin_genes <- c(chromatin_genes, gene)
  }
}
chromatinGenesString <- paste(chromatin_genes, collapse = ", ")
cat(paste("Chromatin Genes: ", chromatinGenesString))
```


