---
title: "MCODE Clustering"
output: html_notebook
---

```{r}
BiocManager::install(c("igraph", "RCy3", "clusterProfiler", "org.Hs.eg.db"))
```
# Get libraries
```{r}
library(RCy3)
library(igraph)
library(clusterProfiler)
library(org.Hs.eg.db)


cytoscapePing ()
```

```{r}
networkSuid = getNetworkSuid()
print(networkSuid)
```

```{r}
commandsHelp("help cluster")
```

```{r}
# minimum number of nodes to create a new network out of the cluster
nodeThreshold <- 15
```
# Perform clustering and save clusters
```{r}
clustermaker = paste("cluster mcode network=SUID:", networkSuid, sep="")
res <- commandsGET(clustermaker)
num <- as.numeric(gsub("Clusters: ", "", res[1]))
print(paste("Found", num, "clusters"))

clusterNumber = 0
# for each cluster
for(i in 1:num) {
  setCurrentNetwork(network=networkSuid)
  clearSelection(network=networkSuid)
  currentCluster <- i
  selectedNodes <- selectNodes(network=networkSuid, currentCluster, by.col="__mcodeCluster") 
  print(paste("Cluster", i, "has", length(selectedNodes$nodes), "nodes"))
  if(length(selectedNodes$nodes) > nodeThreshold) {
    # sub network creation
    # cluster numbers of clusters with #nodes > nodeThreshold
    clusterNumber = clusterNumber + 1
    
    # export subnetwork
    subnetwork_suid <- createSubnetwork(nodes="selected", network=networkSuid)
    exportNetwork(paste("data/networks/MCODE/mcode_cluster_", clusterNumber, ".sif", sep=""), network=subnetwork_suid)
    
    # export nodes of subnetwork
    table <- getTableColumns(table="node",columns = c("display name", "Ensembl", "Entrez Gene"), network = as.numeric(subnetwork_suid))
    write.table(table, file=paste("data/networks/MCODE/mcode_cluster_", clusterNumber,"_nodes.txt", sep=""), row.names=F, sep=",")
    
    res <- clusterProfiler::enrichGO(
      gene = table$`display name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )

    res.df <- as.data.frame(res)
    
    # save enrichment results
    write.table(res.df, file=paste('data/enrichment/molecular-function/mcode_cluster', clusterNumber, '_go.txt', sep=""), row.names=F, sep=",")
    
    assign(paste("MCODE_Cluster_", clusterNumber, sep=""), res.df)
    # plot enrichment results
    barplot(res, showCategory = 10, title = paste("Molecular Function MCODE cluster", clusterNumber))
  }
}
```

```{r}
plotClusterEnrichment <- function(i) {
  tableCluster <- read.csv(paste("data/networks/MCODE/mcode_cluster_", i,"_nodes.txt", sep=""), header=TRUE)
  resCluster <- clusterProfiler::enrichGO(
      gene = tableCluster$`display.name`,
      OrgDb = org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "MF",
      pAdjustMethod ="BH",
      pvalueCutoff = 0.05,
      qvalueCutoff = 0.05,
      readable = TRUE
      )
  resCluster.df <- as.data.frame(resCluster)
  head(resCluster.df)
  barplot(resCluster, showCategory = 15, title = paste("Molecular Function MCODE cluster ",i, sep=""), font.size=9)
}
```


```{r}
plotClusterEnrichment(1)
```

```{r}
plotClusterEnrichment(2)
```

```{r}
plotClusterEnrichment(3)
```

We do not only want over-representation analysis, but to include the order from the log fold change. For this, I need the original list of differentially expressed genes again.


```{r}
i <- 1
# load the gene symbols
tableCluster <- read.csv(paste("data/networks/MCODE/mcode_cluster_", i,"_nodes.txt", sep=""), header=TRUE)
deg_df <- read.csv("data/deg.txt", header=TRUE)
log2FCThreshold <- 0.5
pValueThreshold <- 0.05
deg_cluster <- subset(deg_df, GeneID %in% tableCluster$`Entrez.Gene`)
print(length(deg_cluster$GeneID))
# create the ordered list we need
geneList <- subset(deg_cluster, select=c("GeneID", "log2FoldChange"), padj<pValueThreshold & abs(log2FoldChange) >= log2FCThreshold)
foldChange <- geneList$log2FoldChange
geneListOrdered <- setNames(foldChange, as.character(geneList$GeneID))
geneListOrdered <- sort(geneListOrdered, decreasing=T)
```

With this we can do gene set enrichment (still molecular function)
TODO: does this make sense here? Because we already have a cluster, so over-representation should mean more than this?
```{r}
# for cluster 1, we have only upregulated genes, thats why std scoring does not work
enrichmentGo <- gseGO(geneListOrdered, "org.Hs.eg.db", ont="MF", scoreType="pos")
head(enrichmentGo, 20)
```



